### 2023-04-22 AileensCode.Rmd
a

---
title: "Aileens Code"
author: "Aileen Li"
date: "2023-04-19"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}

```{r delete}
# Use Information Gain for variable selection
library(MASS)
fit_lm = lm(value~, data=train)
stepAIC(fit_lm, direction="backward", trace=FALSE)
stepAIC(fit_lm, direction="forward", trace=FALSE)
stepAIC(fit_lm, direction="both", trace=FALSE)

# Use Recursive Feature Elimination method
set.seed(123)
cparam = rfeControl(functions=lmFuncs, method='repeatedcv', repeats=5, number=10, verbose=FALSE, returnResamp='all')
profile=rfe(x=train_scaled, y=as.matrix(train$value), sizes=c(1:20),rfeControl=cparam)

# Use Lasso for variable selection X
y=train[,2]
library(glmnet)
feat_mod_select = cv.glmnet(train_scaled, train$value, standardize=TRUE, alpha=1)
as.matrix(coef(features,features$lambda.min))

# fit randfom forest and get var. importance based on mean decr. in accuracy
set.seed(123)
library(party)
train_num = train[,2:50] |> dplyr::select(where(is.numeric))
cf1 = cforest(value ~ ., data=train_num, control=cforest_unbiased(mtry=2, ntree=50))
# adj for corr. btw predictors
varimp(cf1, conditional=TRUE)
# more robust towards class imbalance
varimpAUC(cf1)
```

```