### 2023-04-23 Draft.Rmd
Update Draft.Rmd

---
title: "Predicting Ambient Air Pollution (PM2.5) Across the Contiguous U.S."
author: "Aileen Li, Nyah Strickland"
date: "2023-04-18"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Predicting Ambient Air Pollution (PM2.5) Across the Contiguous U.S.

  Air pollution has consequences for everyone, most especially for those with pre-existing conditions. Air pollution is measured in tiny particles or droplets in the air that are two and one-half microns or less in width called PM2.5. So being able to predict the average PM2.5 concentrations ($\mu$g/m$^3$) will allow us to keep people most vulnerable safe as we look for and implement changes to improve the air quality in the region. We created four models to predict the annual average of ambient air pollution in a location based on the variables in the dataset. We chose our predictors using the Random Forest algorithm. And for our four models, we chose Linear Regression, K-Nearest Neighbors, Xgboosting, and Random Forest. Linear Regression models a linear relationship between the PM2.5 value in the atmosphere and the predictors. K-Nearest Neighbors returns predicted PM2.5 values based on their neighboring data points. Xgboosting predicts PM2.5 values by combining the estimates of a set of simpler, weaker models. Random forest makes an averaged prediction from a collection of independent decision trees. We hypothesize that the RMSE performance of the Xgboosting model will be the best, with a value of around 0.5.

```{r Library}
library(tidyverse)
library(tidymodels)
library(broom)
library(xgboost)
library(caret)
library(plotROC)
library(randomForest)
library(factoextra)
library(MASS)
library(ranger)
library(doParallel)

origin <- read_csv("https://github.com/rdpeng/stat322E_public/raw/main/data/pm25_data.csv.gz")
```

## Wrangling

We split 80% of the origin dataset into training and 20% into testing. Then, we scaled the training datasets for PCA and kNN since both of these require little variance among the variables in the dataset. We omitted the 'id' and 'value' variables from the 'trained_scaled' dataset for PCA because the 'id' values for the monitors would not provide any useful insight on PM2.5 concentration and 'value' is not a predictor. Next, we chose 12 principal components with the highest explained variance since the sum of these components met our threshold of 80% explained variance. Then, we tried to see which variables contributed the most to 'value' and reduced the dimension of the 'trained_scale' dataset based on these variables. For kNN, we created a scaled training set 'train_knn' since extreme values would bias the distance calculation used in kNN modeling. The other models either used 'train' dataset since the variables' values do not affect their prediction performance or used a specially altered version of it. Lastly, we did some exploratory analysis to see the linear relationship between longitude and latitude with PM2.5 concentration.

```{r ExploratoryAnalysis}
set.seed(123)
# Split origin dataset into 80/20
dat_split <- initial_split(origin, prop =.8)
dat_split
# Use 80% of data for training 
train <- training(dat_split)
test <- testing(dat_split)

# Create standardize train data for kNN and PCA
# Omit value and id variables since these should not be scaled
# train_scaled for PCA
train_scaled = train[,3:50] %>% dplyr::select(where(is.numeric)) %>% mutate(across(fips:aod, scale)) 
#%>% cbind(state=train$state,county=train$county,city=train$city)

# train_knn for kNN model
train_knn = train[,2:50] %>% dplyr::select(where(is.numeric)) %>% mutate(across(fips:aod, scale)) %>% cbind(state=train$state,county=train$county,city=train$city)

# Calculate principal components on scaled training data
pca_result <- train_scaled %>% dplyr::select(where(is.numeric)) %>%
  prcomp()

# Plot scree plot
pca_result %>% fviz_eig()

# Plot table to show explained variance % by PCs
summary(pca_result)

# Threshold of Explained Variance = 80%
# Calculated the sum of PCs with highest explained variance until threshold was meant
0.3253 + 0.09749 + 0.07984 + 0.06501 + 0.05196 + 0.04147 + 0.03445 +
0.02926 + 0.02563 + 0.0220 + 0.02127 + 0.0188

PC1 <- pca_result$rotation[,1]
PC1_scores <- abs(PC1)
PC1_scores_ordered <- sort(PC1_scores, decreasing = TRUE)
names(PC1_scores_ordered)

# Reduce dimension of dataset
trunc <- pca_result$x[,1:12] %*% t(pca_result$rotation[,1:12])
# Add the center (and re-scale) back to data 
if(all(pca_result$scale != FALSE)){
  trunc <- scale(trunc, center = FALSE , scale=1/pca_result$scale) 
  } 
if(all(pca_result$center != FALSE)){ 
  trunc <- scale(trunc, center = -1 * pca_result$center, scale=FALSE) 
  } 
dim(trunc); dim(train_scaled)

# unhelpful plot -- delete
# plot(pca_result$x[,1], pca_result$x[,2])

# Show which features have the most influence
as_tibble(pca_result$rotation, 
          rownames = "variable") %>% 
    ggplot(aes(variable, PC1)) +
    geom_point() +
    coord_flip()

as_tibble(pca_result$rotation, 
          rownames = "variable") %>% 
    ggplot(aes(variable, PC2)) +
    geom_point() +
    coord_flip()

# Plot clusters between PC1 and PC2 using a biplot
fviz_pca_var(pca_result)

# Reduce dimension of dataset
trunc <- pca_result$x[,1:12] %*% t(pca_result$rotation[,1:12])
# Add the center (and re-scale) back to data 
if(all(pca_result$scale != FALSE)){
  trunc <- scale(trunc, center = FALSE , scale=1/pca_result$scale) 
  } 
if(all(pca_result$center != FALSE)){ 
  trunc <- scale(trunc, center = -1 * pca_result$center, scale=FALSE) 
  } 
dim(trunc); dim(train_scaled)

# Visualize linear relationship between predictors (lat, lon, state) and value
origin %>% ggplot(aes(x=lon, y=value)) + 
  geom_point() + geom_smooth() + 
  labs(title="PM2.5 Concentration vs Longitude", x="Longitude (Degrees)", y="PM2.5 (mu g/m^3)")

origin %>% ggplot(aes(x=lat, y=value)) + 
  geom_point() + geom_smooth() + 
  labs(title="PM2.5 Concentration vs Latitude", x="Latitude (Degrees)", y="PM2.5 (mu g/m^3)")
```

## Results
### Predictor/Feature Extraction

We used the random forest algorithm (RF) to determine which variables would be the best predictors for our model since RF is good at using variance to determine which features would be most influential for a model's predictive performance. In other words, the more variability there is in a dataset, the better RF performs at predicting which variables contribute to the outcome by averaging the results of all trees at the end and this averaging reduces the model's variance. Since our dataset has a wide variety of predictors ranging from education attainment levels to poverty levels in a given monitor region to emission data, RF would be good at reducing the variance in our dataset. For a given feature, the lower its impurity levels are, the more important that feature is. As a result, we got an ordered list of important variables and took the top 10 as input for the training set for our models. The top 10 important predictors were 'CMAQ', 'lat', 'county_area', 'lon', 'aod', 'popdens_county', 'zcta', 'log_nei_2008_pm25_sum_10000', and 'log_nei_2008_pm10_sum_10000'. We will use these features to predict 'value' proceeding onwards. This website was used: [TowardDataScience](https://towardsdatascience.com/feature-selection-using-random-forest-26d7b747597f).

```{r}
# Use Random Forest for variable selection
rfmodel = randomForest(value~., data=train[,2:50])
import=as.data.frame(randomForest::importance(rfmodel)) |> arrange(desc(IncNodePurity))
import=tibble::rownames_to_column(import, 'variables')

# take top 10 important variables
import10 = import[1:10,]
```

### Models 
### 1. Linear Regression
For linear regression (LR), the unscaled training set was used since variance does not affect the prediction performance of this model. Step Akaike Information Criteria (AIC) was used to simplify feature amount without impacting the model's performance. Features are dropped as the AIC score decreases and feature dropping stops once the AIC score increases. 'CMAQ', 'lat', 'lon', 'aod', 'zcta', and 'log_nei_2008_pm25_sum_10000' were determined to be the most important features, and these predictors were used in the recipe. Then, model was fitted. Using the summary function, we found the LR model had a RMSE of 2.22 performance. After 10-fold cross validation, the RMSE score was 2.21, indicating that this model's prediction performance is not good.

```{r LinearRegression}
set.seed(123)
# Fit model and get summary
fit_lm = lm(value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+fips+log_nei_2008_pm10_sum_10000, data=train)
step <- stepAIC(fit_lm, direction="backward", trace=TRUE)
# Drops county_area, popdens_county, fips, and log_nei_2008_pm25_sum_10000

# Create the recipe for all models
rec1 <- train %>% 
    recipe(value ~ CMAQ+lat+lon+aod+zcta+log_nei_2008_pm25_sum_10000)

# Linear regr. model
model1 <- linear_reg() %>% 
    set_engine("lm") %>% 
    set_mode("regression")

wf1 <- workflow() %>% 
    add_recipe(rec1) %>% 
    add_model(model1)
# wf1

res1 <- wf1 |> parsnip::fit(data = train)

# Check performance of linear regression model on the training data
res1 %>% 
    extract_fit_engine() %>% 
    summary()

# Check linear regression model performance using cross-validation
folds1 <- vfold_cv(train, v = 10)
# folds1
res1 <- fit_resamples(wf1, resamples = folds1)
res1 %>% 
    collect_metrics()
```

### 2. K-Nearest Neighbors

The K-Nearest Neighbors (KNN) model finds the predicted 'value' by taking the average 'value' of its k neighbors. We first found the ideal number of k neighbors. This was determined by finding the square root of the number of observations, which was rounded to be 26. We fit the model on the training set data and checked its performance using two methods: the first was fitting the model to extracting the MAE and MSE and the second was 10-fold cross validation. RMSE is the square root of MSE and we were able to confirm its value in the cross-validation by comparing it to the square root of MSE attained from the first method. We found that after 10-fold cross validation, the RMSE was 1.78. This indicates that the model does not predict well since there is such a large amount of error.

```{r kNearestNeighbors}
# kNN uses scaled data called train_knn
set.seed(123)
# Find ideal k neighbors
sqrt(nrow(train))

# Create recipe
rec2 = train_knn |> recipe(value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+fips+log_nei_2008_pm10_sum_10000)

# Create kNN model
model2 <- nearest_neighbor(neighbors = 26) %>% 
    set_engine("kknn") %>% 
    set_mode("regression")

# Create workflow
  wf2 <- workflow() %>% 
    add_recipe(rec2) %>% 
    add_model(model2)

# Fit the model on the training dataset using the `fit()` function
res2 <- parsnip::fit(wf2, data = train_knn)

# Check performance on the complete training data
res2 %>% 
    extract_fit_engine() %>% 
    summary()

# Check performance using cross-validation
folds2 <- vfold_cv(train, v = 10)
# folds
res2 <- fit_resamples(wf2, resamples = folds2) # ERROR Assigned data `orig_rows` must be compatible with existing data.
res2 %>% 
    collect_metrics()
```

### 3. Extreme Gradient Boost Regression

The Extreme Gradient Boosting (XGB) model performs by training weak learners, models that have low prediction accuracy, sequentially to become a single strong learner, a model that has strong prediction accuracy. We preprocess the training set to make the predictive modeling process more accurate and smoother by prepping (estimates the quantities for 'train_scaled') and baking (assigns prepped data to an object) the data. The aforementioned top 10 predictors were used in this preprocessing recipe. Moreover, the categorical variables were factored and features with no variance were removed before this data was baked. Then, this processed data is assigned to 'folds3' and resampled on 5 subsets of the training set. Using 'fold3', we tune the hyperparameters to make the modeling process more efficient and perform grid specification to determine which hyperparameter values have high prediction accuracy (i.e. low prediction error). Next, we isolated the best hyperparameter values, so they can be used for the final boosting model. Lastly, we fitted the model and evaluated its prediction performance. Its RMSE score was 0.56, indicating the XG Boosting is a good model. This website was used: [RBloggers](https://www.r-bloggers.com/2020/05/using-xgboost-with-tidymodels/)

```{r Xgboost}
set.seed(123)
# Preprocessing recipe
preprocessing_recipe <- 
  recipes::recipe(value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+fips+log_nei_2008_pm10_sum_10000, data = training(dat_split)) %>%
  # Convert categorical variables to factors
  recipes::step_string2factor(all_nominal()) %>%
  # Combine low frequency factor levels
  recipes::step_other(all_nominal(), threshold = 0.01) %>%
  # Remove no variance predictors which provide no predictive information 
  recipes::step_nzv(all_nominal()) %>%
  prep()

folds3 <- recipes::bake(preprocessing_recipe, new_data = training(dat_split)) %>%  
  rsample::vfold_cv(v = 5)

# XGBoost model specification
xgboost_model <- parsnip::boost_tree(mode = "regression", trees = 1000, min_n = tune(),
  tree_depth = tune(), learn_rate = tune(), loss_reduction = tune()) %>%
  set_engine("xgboost", objective = "reg:squarederror")

# Grid specification
xgboost_params <- dials::parameters(min_n(), tree_depth(), learn_rate(), loss_reduction())

xgboost_grid <- dials::grid_max_entropy(xgboost_params, size = 60)

head(xgboost_grid)

xgboost_wf <- workflows::workflow() %>%
  add_model(xgboost_model) %>% 
  add_formula(value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+fips+log_nei_2008_pm10_sum_10000)

# Hyperparameter tuning
# Takes a long time to run, 5-8 min
xgboost_tuned <- tune::tune_grid(object = xgboost_wf, resamples = folds3,
  grid = xgboost_grid, metrics = yardstick::metric_set(rmse, rsq),
  control = tune::control_grid(verbose = TRUE))

# Find hyper parameter values which performed best at minimizing RMSE
xgboost_tuned %>% tune::show_best(metric = "rmse") %>%knitr::kable()
# Isolate the best performing hyperparameter values.
xgboost_best_params <- xgboost_tuned %>% tune::select_best("rmse")
xgboost_best_params

# Final boost model
xgboost_model_final <- xgboost_model %>% finalize_model(xgboost_best_params)

# Eval. model performance on train data
train_processed <- bake(preprocessing_recipe,  new_data = training(dat_split))

train_prediction <- xgboost_model_final %>%
  # Fit the model on all the training data
  parsnip::fit(formula = value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+fips+log_nei_2008_pm10_sum_10000, data = train_processed) %>%
  # Predict value for the training data
  predict(new_data = train_processed) %>%
  bind_cols(training(dat_split))

xgboost_score_train <- train_prediction %>%
  yardstick::metrics(value, .pred) %>%
  mutate(.estimate = format(round(.estimate, 2), big.mark = ","))
xgboost_score_train
```

### 4. Random FOrest Regression

The Random Forest (RF) Regression model performs by training multiple different samples simultaneously and averages the outcome at the end for the final prediction. The randomness of these samples contribute to the prediction accuracy of this model. Similar to the XG Boosting Regression model, we preprocess the training set with the same features to improve make the predictive modeling process and then create model specifications. This processed data is assigned to 'folds4' and resampled on 5 subsets of the training set to tune the hyperparameters. We then perform grid specification to determine which hyperparameter values have high prediction accuracy (i.e. low prediction error). Next, we isolated the best hyperparameter values, so they can be used for the final boosting model. Lastly, we fitted the model and evaluated its prediction performance. Its RMSE score was .62, indicating the RF is a good predictive model.

```{r RandomForest}
set.seed(123)
# Preprocessing recipe
rec4 <- recipe(value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+fips+log_nei_2008_pm10_sum_10000, data=train) %>%
  step_log(all_outcomes()) |> 
  step_normalize(all_numeric_predictors()) |> 
  step_dummy(all_nominal_predictors())

#  Create model specification and wf
rf_spec <- rand_forest(mtry = tune(), trees = tune(), min_n = tune()) %>%
  set_engine("ranger", num.threads=7, importance="impurity") |> 
  set_mode("regression")

wf4 <- workflow() %>%
  add_recipe(rec4) %>%
  add_model(rf_spec)

folds4 = recipes::bake(preprocessing_recipe, new_data = training(dat_split)) %>%  
  rsample::vfold_cv(v = 5)

# Space-filling designs grid
rf_grid <- grid_latin_hypercube(min_n(), mtry(range = c(4, 9)), trees(), size = 60)

# Tuning hyperparameters
all_cores = parallel::detectCores(logical=FALSE)
cl = makePSOCKcluster(all_cores)
doParallel::registerDoParallel(cl)

# Takes long time to load 2 min
tune_res <- wf4 |> tune_grid(resamples = folds4, grid = rf_grid, metrics=metric_set(rmse,rsq), control = tune::control_grid(verbose = TRUE))
# tune_res

# Find hyper parameter values which performed best at minimizing RMSE
tune_res %>% tune::show_best(metric = "rmse") %>%knitr::kable()
# Isolate the best performing hyperparameter values.
rf_best_params <- tune_res %>% tune::select_best("rmse")
rf_best_params

# Final Random Forest model
rf_model_final <- rf_spec %>% finalize_model(rf_best_params)

# Check Random Forest model performance
train_processed2 <- bake(preprocessing_recipe,  new_data = training(dat_split))

train_prediction2 = rf_model_final |> 
  parsnip::fit(formula = value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+fips+log_nei_2008_pm10_sum_10000, data=train_processed2) |> 
  predict(new_data=train_processed2) |> 
  bind_cols(training(dat_split))

rf_score_train = train_prediction2 |> 
  yardstick::metrics(value, .pred) |> 
  mutate(.estimate=format(round(.estimate, 2), big.mark=","))
rf_score_train
```

### Best Model

The following code creates a table containing the RMSE and RSQ estimates for each of our four models.

```{r}
# Create table of prediction metrics from training data
linreg = res1 |> collect_metrics() |> as.data.frame()
linreg = cbind(linreg[,1], round(linreg[,3],2))
knn = res2 %>% collect_metrics() |> as.data.frame()
knn = cbind(knn[,1], round(knn[,3],2))
xgboost = xgboost_score_train |> filter(.metric!='mae') |> as.data.frame()
xgboost = cbind(xgboost[,1], xgboost[,3])
ranfor = rf_score_train |> filter(.metric!='mae') |> as.data.frame()
ranfor = cbind(ranfor[,1], ranfor[,3])

compare = c('LR','LR','KNN','KNN','XGB','XGB','RF','RF') |> cbind(rbind(linreg,knn,xgboost,ranfor))
colnames(compare) = c('model','estimator','estimate') 
compare = as.data.frame(compare) |> pivot_wider(names_from=estimator, values_from=estimate)
compare

# Fit RF model to testing data
test_processed <- bake(preprocessing_recipe,  new_data = testing(dat_split))

# Get pred. metrics based on testing data
test_prediction = xgboost_model_final |> 
  parsnip::fit(formula = value ~ CMAQ+lat+county_area+lon+aod+popdens_county+zcta+
             log_nei_2008_pm25_sum_10000+fips+log_nei_2008_pm10_sum_10000, data=test_processed) |> 
  predict(new_data=test_processed) |> 
  bind_cols(testing(dat_split))

xgb_score_test = test_prediction |> 
  yardstick::metrics(value, .pred) |> 
  mutate(.estimate=format(round(.estimate, 2), big.mark=","))
xgb_score_test
```

As seen on the table, LR has the highest RMSE and lowest r-squared estimate. KNN has the second highest RMSE with a moderate R-squared estimate. And XGB has the lowest RMSE of 0.56 and the highest R-squared estimate of 0.96. Random Forest model has the both the second lowest RMSE and second highest R-squared estimate. Since we are choosing the best model by the best RMSE, which is the lowest one, the Xgboost model is thus the best and final model based on our procedure. After fitting the Xgboost model to the testing data, we find that its RMSE and R-squared is 0.61 and 0.94, respectively. Since we are trying to predict annual average ambient air pollution of a given monitor, then these statistics mean that the amount of error in our model's predictions is about 0.61 $\mu$g/m$^3$, and that the features we used in our model are able to explain 94% of the variance in the data, which means this model fits/predicts the data well.

## Results
```{r Results}
compare |> ggplot(aes(y=rmse)) + geom_bar()

```

Your results should include 
at least one visualization demonstrating the prediction performance of your models.

**1. Based on test set performance, at what locations does your model give predictions that are closest and furthest from the observed values? What do you hypothesize are the reasons for the good or bad performance at these locations?**

**2. What variables might predict where your model performs well or not? For example, are their regions of the country where the model does better or worse? Are there variables that are not included in this dataset that you think might improve the model performance if they were included in your model?**

**3. There is interest in developing more cost-effect approaches to monitoring air pollution on the ground. Two candidates for replacing the use of ground-based monitors are numerical models like CMAQ and satellite-based observations such as AOD. How well do CMAQ and AOD predict ground-level concentrations of PM2.5? How does the prediction performance of your model change when CMAQ or AOD are included (or not included) in the model?**

- create visualization showing how close features CMAQ and AOD predict values of PM2.5 to actual/observed values
- no cmaq and no aod models

```{r Q3}

```

**4. The dataset here did not include data from Alaska or Hawaii. Do you think your model will perform well or not in those two states? Explain your reasoning.**

Given that the top 10 features were 'CMAQ,' 'lat,' 'county_area,' 'lon,' 'aod,' 'popdens_county,' 'zcta,' 'log_nei_2008_pm25_sum_10000,' 'fips,' and 'log_nei_2008_pm10_sum_10000,' we think that XG Boost model will perform well since the original dataset included states of varying sizes with different climates and pollution levels. Moreover, its RMSE score is 0.62 and R-squared is 0.94, indicating that the prediction performance of this model is good in relation to the other models. Since Hawaii and Alaska are almost polar opposites with the former being geographically smaller and having hotter climate, we suspect that the model will be able to predict PM2.5 concentration levels in Hawaii and Alaska if they were included in this dataset since the diversity but also inter-relatedness of the features in this dataset aids the model in predicting ambient air pollution in differing environments.

## Discussion
Putting it all together, what did you learn from your data and your model performance?

• Answer your Question posed above, citing any supporting statistics, visualizations, or results from the data or your models.

• Reflect on the process of conducting this project. What was challenging, what have you learned from the process itself?

• Reflect on the performance of your final prediction model. Did it perform as well as you originally expected? If not, why do you think it didn’t perform as well?

• Include acknowledgments for any help received. If a group project, report the contribution of each member (i.e. who did what?).